---
title: "p8105_hw6_mm5351"
author: "Martha Mulugeta"
date: "11/15/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "90%"
  )

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colur_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)
```

**Problem 1**
```{r load/tidy data}
birthweight_data = 
  read_csv("./Data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = recode(babysex, 
              "1" = "male", 
              "2" = "female"),
  babysex = as.factor(babysex),
  frace = recode(frace,
                 "1" = "White",
                 "2" = "Black",
                 "3" = "Asian",
                 "4" = "Puerto Rican",
                 "8" = "Other",
                 "9" = "Unknown"),
  frace = as.factor(frace),
  malform = recode(malform,
                   "0" = "absent",
                   "1" = "present"),
  malform = as.factor(malform),
  mrace = recode(mrace,
                 "1" = "White",
                 "2" = "Black",
                 "3" = "Asian",
                 "4" = "Puerto Rican",
                 "8" = "Other"),
  mrace = as.factor(mrace)) %>% 
  na.omit()
```
The proposed regression model for birthweight includes two predictors: family monthly inclome (fincome) and mother's age at delivery (momage). This model will be referred to as model1. 

```{r model1}
model1 = lm(bwt ~ fincome + momage, data = birthweight_data) 

model1 %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 2)

modelr::add_residuals(birthweight_data, model1) %>% 
modelr::add_predictions(model1) %>% 
ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  labs(
    title = "Model Residuals Against Fitted Values",
    x = "Fitted Values",
    y = "Residuals")
```

The above plot represents the model residuals against fitted values. For the relationship to fit the assumption of linearity, we would expect the residuals to hover around the 0 value. While this is where most of the values exist, outliers are present.

The model previously depicted (model1) will now be compared with two other models for birthweight. The predictors in model2 include baby's length at birth (blength) and gestational age (gaweeks), only looking at the main effects. The predictors in model3 include baby's head circumference (bhead), length at birth (blength), sex (babysex), and all  of the corresponding interactions. 

```{r models}
model2 = 
  lm(bwt ~ blength + gaweeks, data = birthweight_data) %>% 
  broom::tidy()

model3 = 
  lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = birthweight_data) %>% 
  broom::tidy()
```

```{r cv}
cv_df = 
  crossv_mc(birthweight_data, 100)

cv_df %>% pull(train) %>% .[[1]] %>% as_tibble
cv_df %>% pull(test) %>% .[[1]] %>% as_tibble

cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(model1  = map(train, ~lm(bwt ~ fincome + momage, data = birthweight_data)),
         model2  = map(train, ~lm(bwt ~ blength + gaweeks, data = birthweight_data)),
         model3  = map(train, ~ lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = birthweight_data))) %>% 
  mutate(rmse_model1 = map2_dbl(model1, test, ~rmse(model = .x, data = .y)),
         rmse_model2 = map2_dbl(model2, test, ~rmse(model = .x, data = .y)),
         rmse_model3 = map2_dbl(model3, test, ~rmse(model = .x, data = .y)))
```

```{r plot2}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
    geom_violin() +
    labs(
      title = "Model Comparison",
      x = "Model",
      y = "Root Mean Squared Error")
```

The above figure compares the three models based on the root mean squared errors (RMSEs). The RMSEs indicate absolute measure of fit, such that a lower value indicates better fit. Based on this, it is evident that model3 is the best fitting model, followed by model2, and then model1. 

***Problem 2***
```{r load/tidy}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r rsquared}
bootstrap_rsquared = 
  weather_df %>%
  modelr::bootstrap(n = 5000) %>%
  mutate(
    models = map(strap, ~ lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::glance)) %>% 
  select(results) %>% 
  unnest(results) %>% 
  mutate(
    lower = quantile(r.squared, 0.025),
    upper = quantile(r.squared, 0.975),
    lower = round(lower, digits = 2),
    upper = round(upper, digits = 2)
  ) %>% 
  select(lower, upper) %>% 
  distinct()
```

The 95% confidence interval for r-squared is between `r pull(bootstrap_rsquared, lower)` and `r pull(bootstrap_rsquared, upper)`.

```{r log}
bootstrap_log = 
  weather_df %>%
  modelr::bootstrap(n = 5000) %>%
  mutate(
    models = map(strap, ~ lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy)) %>% 
  select(results) %>% 
  unnest(results) %>%
  select(term, estimate) %>% 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) %>% 
  unnest() %>% 
  rename("intercept" = "(Intercept)") %>% 
  mutate(logb0b1 = log(intercept * tmin),
    lower = quantile(logb0b1, 0.025),
    upper = quantile(logb0b1, 0.975), 
    lower = round(lower, digits = 2),
    upper = round(upper, digits = 2)
  ) %>% 
  select(lower, upper) %>% 
  distinct()
```

The 95% confidence interval for log(b0 * b1) is between `r pull(bootstrap_log, lower)` and `r pull(bootstrap_log, upper)`.


